---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library("knitr")
activity <- read.csv("activity.csv")
activity$date <- as.Date(as.character(activity$date),"%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r}
stepsDaily<- with(activity,tapply(activity$steps,activity$date,sum))
hist(stepsDaily)
```

Mean steps by day
```{r}
data.frame(with(activity,tapply(activity$steps,activity$date,mean)))
```

Median steps by day
```{r}
data.frame(with(activity,tapply(activity$steps,activity$date,median)))
```

## What is the average daily activity pattern?

Average number of steps taken, averaged across all days (y-axis)
```{r}
byMinuteInterval <- with(activity,tapply(activity$steps,activity$interval,mean,na.rm=TRUE))
plot(byMinuteInterval,type='l')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max(byMinuteInterval)
```

Knowing the maximum value, we look for the minute interval.
```{r}
maxByMinute<-byMinuteInterval[byMinuteInterval>206]
```

The minute interval is `r names(maxByMinute)`.

## Imputing missing values

```{r}
total_nas <- sum(is.na(activity[,1]))
```
The total number NA steps data is `r total_nas`

Filling `r NA` values with the *mean* by minute interval.

```{r}
for (i in 1:length(activity$steps)) {
    if(is.na(activity$steps[i])){
        boolInterval <- activity$interval[i]==as.numeric(as.character(names(byMinuteInterval)));
        activityFilled[i,1]<-byMinuteInterval[boolInterval]
    } else {
        activityFilled[i,1]<-activity$steps[i]
    }
    if(exists("boolInterval")) rm(boolInterval)
}
```

Creating a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activityFilled2<-data.frame()
activityFilled2<-cbind(as.double(activityFilled[,1]),as.character(activity$date),as.integer(activity$interval))
colnames(activityFilled2)<-colnames(activity)
activityFilled2<-as.data.frame(activityFilled2)
```

Histogram of steps by day with the filled data.

```{r}
activityFilled2$date <- as.Date(as.character(activityFilled2$date),"%Y-%m-%d")
rm(stepsDaily)
stepsDaily<- with(activityFilled2,tapply(as.numeric(activityFilled2$steps),activityFilled2$date,sum))
hist(stepsDaily)
```

Mean steps by day with the filled data.

```{r}
data.frame(with(activityFilled2,tapply(as.numeric(activityFilled2$steps),activityFilled2$date,mean)))
```

Median steps by day with the filled data.

```{r}
data.frame(with(activityFilled2,tapply(as.numeric(activityFilled2$steps),activityFilled2$date,median)))
```
RJOF: Do these values differ from the estimates from the first part of the assignment?

RJOF: What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?

New variable indicating if is *weekday* or *weekend*

```{r}
activityFilled2$w <- ifelse((weekdays(activityFilled2$date) == "Saturday") | (weekdays(activityFilled2$date) == "Sunday"), "weekend", "weekday")
```

Subseting for *weekdays* and *weekends*

```{r}
activityWeekend <- activityFilled2[activityFilled2$w=='weekend',]
activityWeekday <- activityFilled2[activityFilled2$w=='weekday',]
```

Generating the mean by minutes interval.

```{r}
byMinuteIntervalWeekend <- with(activityWeekend,tapply(as.numeric(activityWeekend$steps),activityWeekend$interval,mean))

byMinuteIntervalWeekday <- with(activityWeekday,tapply(as.numeric(activityWeekday$steps),activityWeekday$interval,mean))
```

Ploting.

```{r}
par(mfcol=c(2,1))
plot(byMinuteIntervalWeekend,type='l',xlab='weekend')
plot(byMinuteIntervalWeekday,type='l',xlab='weekday')
```